---
title: "Group_5_Analysis"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)
library(tidyverse)
library(gt)
library(MASS)
library(patchwork)
library(gapminder)
library(janitor)
library(tidymodels)
library(moderndive)
library(ggstats)
```

# Introduction {#sec-intro}
Obesity levels across Scotland are monitored as part of the Scottish Health Surveys. Within these surveys there is information on socio-economic and lifestyle factors and their effect on the prevalence of obesity in Scotland. Using data gathered between 2013 and 2016, this report will look at whether or not the prevalence of obesity has changed over this time period. Then, in addition, looking at whether there are differences in obesity prevalence caused by age, sex and socio-economic status.

```{r}
#| echo: false

# Read in the data
Project11 <- read.csv("~/Desktop/STATS4052/DAS-Group-5/DAProject11.csv")

# Add a new column called IsObese based on the value of BMI group
Project11 <- Project11 |>
  dplyr::mutate(IsObese = ifelse(BMIgroup == "Obese", "Obesity", "Not Obesity"))
```

# Exploratory Analysis {#sec-exp}

```{r}
#| echo: false
#| label: tbl-proportion
#| tbl-cap: "Proportion of obseity by year"

Project11 |>
  tabyl(IsObese,Year) |>
  adorn_percentages() |>
  adorn_pct_formatting() |>
  adorn_ns() |>
  gt()
```

```{r}
#| echo: false
#| label: fig-barplot
#| message: false
#| fig-cap: "Barplot of year by Obesity"
#| fig-width: 4
#| fig-height: 3
#| fig-align: center


ggplot(data = Project11, aes(x = Year, group = IsObese))+
  geom_bar(aes(y = after_stat(prop), fill = IsObese), stat = "count", position = "dodge") +
  labs(y = "Proportion", fill = "IsObese")
```

```{r}
#| echo: false
#| label: tble-logodds
#| message: false
#| tbl-cap: "Summary of logistic regression model"



```

# Formal Analysis {#sec-formal}

# Conclusion {#sec-con}
